---
title: "How do environmental stressors and resource availability influence reproductive success of an urban adapring songbird?"
author: Me
date: today
toc: true
abstract-titel: "Our Hypothesis:"
abstract: "With higher insect abundance and lower noise levels we expect to encounter more fledglings. The variables have an interaction which inluences the fledgling numbers differently"

project:
  output-dir: output
  execute-dir: project

format:
  html:
    embed-resources: true
    number-sections: true
    code-fold: true
    comments: 
      hypothesis: true
    
editor: visual
editor_options:
  chunk_output_types: console

execute:
  warning: false
  error: false
---

# Preparing session

```{r}
#| label: preparing the data
#| output: false
#| echo: false

# rm(list=ls())

# 0. Load libraries


library(ggplot2)
library(pastecs)
library(ggpubr); theme_set(theme_bw())
library(GGally)
library(performance)
library(see)
library(DHARMa)
library(car)
library(ggeffects)
library(MuMIn) #rsquaredLR is from here
library(knitr)
library(dplyr)
library(readr)

# 1. Import and correct data
urb_fledg <- read_csv(
  "data/processed/UrbanFledgling.csv",
  col_types = "dddc"
)

#remove the 2 faulty rows
urb_fledg<-urb_fledg%>%
  filter(is.na(urb_fledg$Comments))

urb_fledg<-urb_fledg%>%
  select(-Comments)

str(urb_fledg) #insect abundance seems to be a chr
summary(urb_fledg)

```

## Explore data - Numerically

```{r}
#| label: exp-numerical

cor_insect_vs_noise<-cor(urb_fledg$insect_abundance_g_m_2, urb_fledg$environmental_noise_dB)
#strong negative correlation

```

We have a negative correlation of `r round(cor_insect_vs_noise, 2)` which means that there is an interaction between the environmental noise and insect abundance.

## Explore data - Graphically

```{r}
#| label: fig-exp-graph
#| fig-cap: "Exploratory plots; a) higher environmental noise (dB) is associated with fewer fledglings; b) higher insect abundance (g/m^2) is associated with a greater number of fledglings."
#| fig-width: !expr 15/2.54
#| fig-height: !expr 7/2.54

ggarrange(
  ggplot(urb_fledg, aes(x=environmental_noise_dB, y=number_of_fledglings))+
    geom_point()+
    geom_smooth(method="lm")+
    xlab("Enrironmental noise (dB)")+
    ylab("Number of fledglings")
  ,
  
  ggplot(urb_fledg, aes(x=insect_abundance_g_m_2, y=number_of_fledglings))+
    geom_point()+
    geom_smooth(method="lm")+
    xlab("Insect abundance (g/m^2)")+
    ylab("Number of fledglings"),
  common.legend = TRUE, labels=c("a)", "b)")
  
)

```

As we can see in the @fig-exp-graph we have a positive correlation between the number of fledglings and the insect abundance and a negative correlation to environmental noise.


# Fit the models

## Fit candidate models
```{r}
#| label: tbl-candidate-models
#| tbl-cap: "Candidate model comparisons"

model.00<-glm(number_of_fledglings~1, urb_fledg, family=poisson)
model.01<-glm(number_of_fledglings~insect_abundance_g_m_2, urb_fledg, family=poisson)
model.02<-glm(number_of_fledglings~insect_abundance_g_m_2+environmental_noise_dB, urb_fledg, family=poisson)
model.03<-glm(number_of_fledglings~insect_abundance_g_m_2*environmental_noise_dB, urb_fledg, family=poisson)

knitr::kable(compare_performance(model.00, model.01, model.02, model.03), digits=3)

```

























```{r}


ggplot(urb_fledg, aes(x=insect_abundance_g_m_2, y=environmental_noise_dB))+
  geom_point(aes(size=number_of_fledglings, color=number_of_fledglings), alpha=.7)
```







